---
title: "Sampling Distribution"
format: html
warning: false
---

**Sampling Distributions and the Central Limit Theorem**  

When estimating population parameters like $\mu$, we use sample statistics $\bar{X}$. Different samples of size $n$ yield different means - these sample means are random variables with their own distribution. The **sampling distribution** describes all possible values of $\bar{X}$ from repeated sampling. Key properties:

1. **Unbiasedness**: $\mu_{\bar{X}} = \mu$  
2. **Precision**: $\sigma_{\bar{X}} = \sigma/\sqrt{n}$  

The Central Limit Theorem (CLT) reveals the shape:  
- **Exact normality**: If population is normal (any $n$)  
- **Approximate normality**: For any population with $n \geq 30$  

```{r}
#| echo: false
# Population and sampling distribution simulation
library(ggplot2)
set.seed(123)
population <- rexp(100000, 0.5)  # Skewed population (μ=2)
n <- 50
sample_means <- replicate(5000, mean(sample(population, n)))
```

**Population Distribution**  

```{r}
#| echo: false
# Population plot
ggplot() +
  geom_histogram(aes(x = population, y = after_stat(density)), 
                 bins = 50, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = 2, color = "red", linewidth = 1) +
  annotate("text", x = 6, y = 0.3, 
           label = "Exponential Population\n(μ = 2)", color = "steelblue") +
  labs(title = "Parent Population Distribution",
       x = "Value", y = "Density") +
  theme_minimal()
```


*Exponential distribution with population mean*

**Sampling Distribution**  

```{r}
#| echo: false
# Sampling distribution plot
ggplot() +
  geom_histogram(aes(x = sample_means, y = after_stat(density)), 
                 bins = 50, fill = "darkgreen", alpha = 0.7) +
  stat_function(fun = dnorm, color = "red", linewidth = 1,
                args = list(mean = 2, sd = sd(population)/sqrt(n))) +
  annotate("text", x = 2.8, y = 1.2, 
           label = "CLT Normal \n(μ = 2, SE = σ/√n)", 
           color = "red") +
  labs(title = "Sampling Distribution of Sample Means (n = 50)",
       x = expression(bar(X)), y = "Density") +
  theme_minimal()
```

 
*CLT convergence despite non-normal parent distribution*

**Key Observations**  

- Sample means cluster around population mean (unbiasedness)  
- Spread decreases with larger $n$ (precision improvement)  
- Normal shape emerges despite skewed population (CLT in action)  

**Critical Implications**  

- Enables inference even with unknown population distributions  
- Justifies common statistical procedures (confidence intervals, hypothesis tests)  
- Fails if:  

  ‣ Small samples ($n < 30$) from non-normal populations  
  ‣ Population variance is infinite  
  
  
